---
title: "Robust Tests"
author: "Qi"
date: "`r Sys.Date()`"
format: html
execute: 
  warning: false
  message: false
---

```{r sandwich}
# Load necessary packages
# install.packages(c("sandwich", "car", "boot", "robust", "robustbase", "caret", "plm"))
library(sandwich)
library(car)
library(boot)
library(robust)
library(robustbase)
library(caret)
library(plm)
library(lmtest)
```

## Simulate the dataset

```{r simulation}
set.seed(123)
n <- 100
x1 <- rnorm(n, mean = 5, sd = 2)
x2 <- rnorm(n, mean = 10, sd = 3)
y <- 3 + 2*x1 + 1.5*x2 + rnorm(n, mean = 0, sd = 2)
mydata <- data.frame(y, x1, x2)
head(mydata)
```

## Example with sandwich package (Robust Standard Errors)

```{r}
lm_model <- lm(y ~ x1 + x2, data = mydata)
robust_se <- coeftest(lm_model, vcov = vcovHC(lm_model, type = "HC1"))
print("Robust Standard Errors using sandwich package")
print(robust_se)
```

## Example with car package (VIF)

```{r vif}
vif_values <- vif(lm_model)
print("Variance Inflation Factor using car package")
print(vif_values)
```

## Example with boot package (Bootstrapping)

```{r bootstrap}
boot_function <- function(data, indices) {
  d <- data[indices, ]
  lm_model <- lm(y ~ x1 + x2, data = d)
  return(coef(lm_model))
}
boot_results <- boot(data = mydata, statistic = boot_function, R = 1000)
print("Bootstrapping using boot package")
print(boot_results)
```

## Example with robust package (Robust Linear Model)

```{r robust_linear}
rlm_model_robust <- robust::lmRob(y ~ x1 + x2, data = mydata)
print("Robust Linear Model using robust package")
print(summary(rlm_model_robust))
```
## Example with robustbase package (Robust Linear Model)
```{r robustbase}
rlm_model_robustbase <- robustbase::lmrob(y ~ x1 + x2, data = mydata)
print("Robust Linear Model using robustbase package")
print(summary(rlm_model_robustbase))
```
## Example with caret package (Cross-Validation)

```{r caret}
train_control <- trainControl(method = "cv", number = 10)
model_caret <- train(y ~ x1 + x2, data = mydata, method = "lm", trControl = train_control)
print("Cross-Validation using caret package")
print(model_caret)
summary(model_caret)
```

## Example with plm package (Panel Data Analysis with Robust Standard Errors)

_Note_: For plm, we need to simulate panel data. We'll create a simple panel data example.

```{r plm}
set.seed(123)
id <- rep(1:20, each = 5)
time <- rep(1:5, times = 20)
x1_panel <- rnorm(100, mean = 5, sd = 2)
x2_panel <- rnorm(100, mean = 10, sd = 3)
y_panel <- 3 + 2*x1_panel + 1.5*x2_panel + rnorm(100, mean = 0, sd = 2)
panel_data <- data.frame(id, time, y = y_panel, x1 = x1_panel, x2 = x2_panel)
head(panel_data)
```



```{r panel1}
# Transform the data into a pdata.frame
panel_data <- pdata.frame(panel_data, index = c("id", "time"))

# Fit a panel model and calculate robust standard errors
plm_model <- plm(y ~ x1 + x2, data = panel_data, model = "within")
summary_plm <- summary(plm_model, vcovHC(plm_model, method = "arellano"))
print("Panel Data Model with Robust Standard Errors using plm package")
print(summary_plm)
```
