---
title: "Causal Inference matching"
format: html
---

## Intro

This is one of my assignment of the course _A Crash Course in Causality: Inferring Causal Effects from Observational Data_ on Coursera. Module 3, Matching.

## Running Code

```{r loadLib}
library(tableone)
library(Matching)
library(MatchIt)
data(lalonde)
```
```{r}
summary(lalonde)
```

To calculate the standardized mean difference.

```{r sme}
#| echo: false

xvars = c('age', 'educ', 'black', 'hispan', 'married', 'nodegree', 're74', 're75')
#look at a table 1
table1<- CreateTableOne(vars=xvars,strata="treat", data=lalonde, test=FALSE)
## include standardized mean difference (SMD)
print(table1,smd=TRUE)
```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
library(dplyr)

result <- lalonde %>%
  group_by(treat) %>%
  summarize(mean_re78 = mean(re78, na.rm = TRUE))

# Print the result
print(result)

result[2, 2] - result[1, 2]
```

## Fit a propensity score

```{r propensityScore}
mydata <- lalonde %>% 
  mutate(black = race == 'black'
         , hispan = race == 'hispan')
psmodel<-glm(treat~age+educ+race+hispan+married+nodegree+re74+re75
             , family=binomial()
             , data=mydata)

#show coefficients etc
summary(psmodel)
#create propensity score
pscore<-psmodel$fitted.values

summary(pscore)
```

## Propensity Score Matching

```{r pmatching}
set.seed(931139)

psmatch<-Match(Tr=mydata$treat, M=1,X=pscore,replace=FALSE)

names(psmatch)

matched<-mydata[unlist(psmatch[c("index.treated","index.control")]), ]

head(matched)

#get standardized differences
matchedtab1<-CreateTableOne(vars=xvars
                            , strata="treat"
                            , data=matched
                            , test = FALSE)
print(matchedtab1, smd = TRUE)
```

## Matching with a Caliper 

```{r pmatchCaliper}
set.seed(931139)

psmatch_c <-Match(Tr=mydata$treat
               , M=1
               , X=pscore
               , replace=FALSE
               , caliper = 0.1)

matched_c <- mydata[unlist(psmatch_c[c("index.treated","index.control")]), ]

head(matched_c)
```

```{r smd_c}
#get standardized differences
matchedtab1_c <- CreateTableOne(vars=xvars
                            , strata="treat"
                            , data=matched_c
                            , test = FALSE)
print(matchedtab1_c, smd = TRUE)
```

```{r}
result <- matched_c %>%
  group_by(treat)%>%
  summarize(mean_re78 = mean(re78, na.rm = TRUE))

# Print the result
print(result)

result[2, 2] - result[1, 2]
```

## Paired t test
```{r paired_t_test}
#outcome analysis
y_trt<-matched_c$re78[matched_c$treat==1]
y_con<-matched_c$re78[matched_c$treat==0]

#pairwise difference
diffy<-y_trt-y_con

#paired t-test
t.test(diffy)
```


